We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-02-21 paper/paper.tex
edit
new intro

The spatial distribution of satellite galaxies around our Milky Way
(MW) and the M31 galaxy provides a stringent test for structure
formation theories in a explicit cosmological context
\citep{2015ApJ...815...19P, 2016MNRAS.457.1931S}.
The presence of a Vast Polar Structure of satellites around the Milky
Way, first suggested by \cite{1976RGOB..182..241K} and
\cite{1976MNRAS.174..695L} as a Magellanic Plane, has been
stablished in the last decade \citep{2005A&A...431..517K,2013MNRAS.435.1928P}. 
This structure is present in the $11$ brightest "classical" dwarf
galaxies and has been confirmed as fainter satellites have been
detected \citep{2014MNRAS.441.2124B}. 
In the case of M31, observational studies have found a planar satellite
distribution in a subset of $15$ satellites out of the 
total poulation of $27$ satellites.\citep{2013ApJ...766..120C,
2013Natur.493...62I}, although its brightest satellites
do not show any special feature and are consistent with a random
spherically symmetric distribution \citep{2006AJ....131.1405K,
  2007MNRAS.374.1125M, 2013ApJ...766..120C}.

>>>>

The spatial distribution of satellite galaxies around our Milky Way
(MW) and the M31 galaxy is becoming a stringent test for structure
formation theories in a explicit cosmological context. 
This started with the suggestion of a Magellanic Plane, a flattened
structure of satellite galaxies around the MW, first suggested by
\cite{1976RGOB..182..241K} and \cite{1976MNRAS.174..695L}.

\cite{2005A&A...431..517K} then quantified that the
highly planar distribution of the 11 classical MW satellites has less
than a $0.5\%$ chance  to happen by chance if the
parent distribution is spherically symmetric, interpreting this as a
challenge to the Lambda Cold Dark Matter (LCDM) paradigm.
The same year it was recognized, using numerical simulations, that the
satellites should not be compared with an spherically
symmetric distribution given that Dark Matter halos in LCDM
are expected to be triaxial
\citep{2005MNRAS.363..146L, 2005ApJ...629..219Z}.
Nevertheless, estimates of the chances to find simulated satellites as
planar as in the MW were inconclusive. 
\cite{2005MNRAS.363..146L} reported that
satellite planes should be expected in every DM while
\cite{2005ApJ...629..219Z} gave an estimate of a $2\%$ probability to
have an arrangement as extreme as in observations.

Later, \cite{2007MNRAS.374.1125M} used simulations to confirm the low chances
found in \cite{2005A&A...431..517K}.
This result was challenged by \cite{2009MNRAS.399..550L} who continue
to report a high chances to find a planar configuration;
the most recent numerical experiment with high resolution simulations
by \cite{2013MNRAS.429..725S} reports low chances to find the
observational result, but precise figures are not quoted in either of
these two reports. 
\citep{2015ApJ...815...19P, 2016MNRAS.457.1931S}.
The presence of a Vast Polar Structure of satellites around the Milky
Way,
stablished in the last decade
\citep{2013MNRAS.435.1928P}. 
This structure is present in the $11$ brightest "classical" dwarf
galaxies and has been confirmed as fainter satellites have been
detected \citep{2014MNRAS.441.2124B}. 
In the case of M31, observational studies have found a planar satellite
distribution in a subset of $15$ satellites out of the 
total poulation of $27$ satellites.\citep{2013ApJ...766..120C,
2013Natur.493...62I}, although its brightest satellites
do not show any special feature and are consistent with a random
spherically symmetric distribution \citep{2006AJ....131.1405K,
  2007MNRAS.374.1125M, 2013ApJ...766..120C}.
